The expected shape of the Milky Way's Dark Matter halo

### 2018-09-18 Paper/paper.tex
edit

A robust prediction of the Cold Dark Matter (CDM) model is that DM
halos are ellipsoidal and can be characterized by the principal axes
$a>b>c$.
The ellipsoidal shape is mostly due to the anisotropical and
clumpy accretion of matter influenced by environmental structures. 
The shape has a strong mass dependence \citep{Allgood_et_al._2006}. 
The general consensus from numerical studies is that Milky Way type
halos tend to be prolate with axial ratios $b/a\approx X$ and  $c/a\approx X$, 
Halos are also found to be rounder at the outerskirts and at lower
redshifts.

@@

Better connector with observations of distant galaxies. Gravitational
lensing and star tracers. (Talk about discrepancy with simulations,
perhaps later) and state that the preference for axisymmetrical shapes
may arise from the observational restrictions (projection) look for
citations about this .\\

In the case of the Milky Way, where we have a three-dimensional view from the inside, there seems to be a general agreement for oblate (i.e. a=b>c) configurations at small distances around \~ <20kpc \citep[see][]{LM10,Bovy2016,Loebman,Vera-Ciro, RobOllingMerrifield,Bovy,BanerjeeChanda}, and more triaxial and prolate configurations on the outter distances \~>20kpc \citep[see][]{Vera-Ciro,LMJ09,Deg,BanerjeeChanda}. However, some studies are inclined towards prolate configurations even at the inner parts of the halo \citep[see][]{BowdenEvansWilliams,Findothers}, and although it previously seemed that a triaxial DM halo on the outerskirts would be necessary to fully explain the characterization of the Sagittarius stream \citep[see][]{LMJ09}, more recent studies put into doubt (question) this claim by reporting inconsistencies with narrow stellar streams \citet{PearsonKupperJohtnston, MoreLikeThis} or finding that the relaxation of other constraints may make this claim unnecessary (wording) \citet{IbataLewisMartin}.\\

>>>>

A robust prediction of the Cold Dark Matter (CDM) paradigm is that DM
halos are ellipsoidal and can be characterized by the principal axes
$a>b>c$.
This ellipsoidal shape is mostly due to the anisotropical and
clumpy accretion of matter influenced by environmental structures. 
Numerical studies how that the shape has a strong mass dependence
\citep{Allgood_et_al._2006}, halos are also rounder at the outerskirts
than at the inner part. Shape also evolves with cosmic time, halos get
rounder as they evolve.

In the case of the Milky Way, where we have a three-dimensional view
from the inside, we do not have a clear agreement. 
Some studies prefer oblate (i.e. a=b>c) configurations at small
distances around $\leq 20$ kpc \citep[see][]{Law_and_Majewski_2010,Bovy_et_el._2016,Loebman_et_al._2012,Olling_and_Merrifield_2000,Banerjee_and_Chanda_2011}
and more triaxial and prolate configurations on the outter distances
$\geq 20$ kpc 
\citep[see][]{Vera-Ciro_and_Helmi_2013,Law_and_Majewski_2009,Deg_and_Widrow_2013,Banerjee_and_Chanda_2011}.
However,
some 
studies are inclined towards prolate configurations even at the inner
parts of the halo \citep[see][]{Bowden_et_al._2016}, and
although it previously seemed that a triaxial DM halo on the
outerskirts would be necessary to fully explain the characterization
of the Sagittarius stream \citep[see][]{LMJ09}, more recent studies
put into doubt (question) this claim by reporting inconsistencies with
narrow stellar streams \citet{PearsonKupperJohtnston, MoreLikeThis} or
finding that the relaxation of other constraints may make this claim
unnecessary (wording) \citet{IbataLewisMartin}.

%Better connector with observations of distant galaxies. Gravitational
%lensing and star tracers. (Talk about discrepancy with simulations,
%perhaps later) and state that the preference for axisymmetrical shapes
%may arise from the observational restrictions (projection) look for
%citations about this.

@@
